### 2021-03-08 text/draft.tex
machine indexing

Annif vocs that are available

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Extract the keywords from the sample data set.
\item
  Clean the extracted keywords.
\item
  Reconcile the extracted keywords with.
\end{enumerate}

The intended output of this transformation process is\ldots{}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords }\OperatorTok{=}\NormalTok{ Keywords.extract\_keywords(DIR }\OperatorTok{+} \StringTok{"/sample/sample\_master.json"}\NormalTok{)}
\NormalTok{Utility.save\_json(keywords, DIR }\OperatorTok{+} \StringTok{"/keywords/keywords\_extracted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_extracted }\OperatorTok{=}\NormalTok{ Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/keywords/keywords\_extracted.json"}\NormalTok{)}
\NormalTok{keywords\_clean }\OperatorTok{=}\NormalTok{ Keywords.clean\_keywords(keywords\_extracted)}
\NormalTok{Utility.save\_json(keywords\_clean, DIR }\OperatorTok{+} \StringTok{"keywords/keywords\_clean.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_clean }\OperatorTok{=}\NormalTok{ Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/keywords/keywords\_clean.json"}\NormalTok{)}
\NormalTok{keywords\_histogram }\OperatorTok{=}\NormalTok{ Keywords.make\_histogram(keywords\_clean)}
\NormalTok{Utility.save\_json(keywords\_histogram, DIR }\OperatorTok{+} \StringTok{"keywords/keywords\_clean\_histogram.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_data\_set }\OperatorTok{=}\NormalTok{ Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master.json"}\NormalTok{)}
\NormalTok{keywords\_counted }\OperatorTok{=}\NormalTok{ Keywords.make\_count(sample\_data\_set)}
\NormalTok{Utility.save\_json(keywords\_counted, DIR }\OperatorTok{+} \StringTok{"/analysis/keywords\_counted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The median number of keywords for an item in the sample data set is 6
with an IQR of 4 (min = 0, Q1 = 4, M = 6, Q3 = 8, max = 87). Of course,
items in the sample data set with a number of keywords below the first
and above the third quartile are highly suspect from a qualitative point
of view: when it comes to subject indexing, some terms are required, but
more is usually worse {[}!! source{]}. Items with too few or too many
keywords will be discussed in more detail in section
\protect\hyperlink{assessment}{section Assessment}

@@

As explained in section X, Annif assigns index terms from a controlled
vocabulary. If we want to assess the quality of the indexing via a gold
standard, we must therefore ensure that the gold standard makes use of
the vocabulary used by Annif. As seen in section Y, the relevant
(English) vocabularies are Wikidata YSO. The next step in constructing
the derivative gold standard is hence to match the extracted and cleaned
keywords with keywords from YSO and Wikidata. This process is called
``reconciliation'' (see https://docs.openrefine.org/manual/reconciling)
and the tool of choice for this task is OpenRefine (see
\protect\hyperlink{openrefine}{section OpenRefine}).

In this section, I will describe how the cleaned keywords were
reconciled with Wikidata and YSO, and which additional steps for
refinement were undertaken. In total, 2'104 data transformation
operations were performed; the complete operation history is available
as \texttt{/keywords/operation\_history.json}.

@@

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API
for details). For example, {[}keyword/article{]}. To generalize,
suggestions with types whose concepts are proper names are usually
incorrect. Based on this observation, suggestions with the following
types were rejected: ``scholarly article,'' ``clinical trial,''
``scientific journal,'' ``academic journal,'' ``open access journal,''
``thesis,'' ``doctoral thesis,'' ``natural number.'' Suggestions with
the following types were manually verified (i.e., checked for
correctness): ``human,'' ``album,'' ``film,'' ``musical group,''
``business,'' ``literary work,'' ``television series,''
``organization,'' ``family name,'' ``written work,'' ``video game,''
``single, television series episode,'' ``painting,'' ``commune of
France,'' ``city of the United States,'' ``magazine,'' ``studio album,''
``year,'' ``nonprofit organization,'' ``border town,'' ``international
organization,'' ``political party,'' ``software,'' ``song,''
``website,'' ``article comic strip,'' ``collection,'' ``commune of
Italy,'' ``fictional human,'' ``film,'' ``government agency,''
``village,'' ``academic journal article,'' ``female given name,''
``poem.''

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_reference }\OperatorTok{=}\NormalTok{ Utility.load\_json(DIR }\OperatorTok{+} \StringTok{"/keywords/keywords\_reference.json"}\NormalTok{)}
\NormalTok{keywords\_reference\_new }\OperatorTok{=}\NormalTok{ Keywords.enrich\_with\_yso(keywords\_reference)}
\NormalTok{Utility.save\_json(keywords\_histogram, DIR }\OperatorTok{+} \StringTok{"keywords/keywords\_reference\_master.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

In this section I will assess the quality of the sample data set's
indexing with Annif based on the gold standard.

@@

Let us briefly look at the definitions and motivations of the chosen
metrics. Remember that a suggestion of a subject term is correct if and
only if the subject term is in the derivative gold standard. The
possible outcomes are summarized in Table 1.

@@

The above scoring metrics were implemented using the scikit-learn
machine learning library (Pedregosa et al. 2011). For a given Annif
configuration
(\protect\hyperlink{annif-versus-derivative-gold-standard}{section
``Annif versus derivative gold standard''})\ldots{} {[}perhaps explain
how it is done, important to mention transformation of multilabel to
binary{]}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Analysis.super\_make\_metrics(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}
\centering
\includegraphics{images/metrics_all_n.pdf}
\caption{Distribution of weighted F1-scores per maximum number of
suggestions 1 \(\leq\) n \(\leq\) 10. See Table 2 for numeric values. n
= 2 is the best perfomring type of configuration. However, it is
outperformed by n = 4 and n = 5 token configurations. The distribution
of the distributions is skewed left. Types of configurations with a
smaller maximum number of suggestions display a more consistent
performance.}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{for}\NormalTok{ dept }\KeywordTok{in}\NormalTok{ Data.get\_departments():}
\NormalTok{    Analysis.super\_make\_metrics(DIR }\OperatorTok{+} \StringTok{"/indexed/indexed\_master.json"}\NormalTok{, dept)}
\end{Highlighting}
\end{Shaded}

>>>>

The vocabularies provided by the Annif API are the General Finnish
Ontology or YSO in short (see https://finto.fi/yso/en/) and Wikidata
(see https://www.wikidata.org/). Each vocabulary is combined with an
algorithm or algorithm ensemble respectively and a specific language to
create a so-called Annif ``project.'' The five projects used in what
follows are summarized in Table 1. A high-level description of the
algorithm(s) is given by Suominen (2019, 7--10). Information on
implementation can be found at the Annif GitHub Wiki (see
https://github.com/NatLibFi/Annif/wiki).

\begin{table}[]
\centering
\begin{tabular}{llll}
Annif project   & Vocabulary & Language & Algorithm(s)   \\ \hline
\texttt{yso-en}  & YSO        & English  & NN ensemble \\
\texttt{yso-maui-en}     & YSO        & English  & Maui        \\
\texttt{yso-bonsai-en}   & YSO        & English  & Omikuji     \\
\texttt{yso-fasttext-en} & YSO        & English  & fastText    \\
\texttt{wikidata-en}     & Wikidata   & English  & TF-IDF     
\end{tabular}
\caption{Annif projects used to index the items of the Edoc sample data set.}
\label{tab:my-table}
\end{table}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Extract the keywords from the sample data set.
\item
  Clean the extracted keywords.
\item
  Reconcile the extracted keywords with Wikidata and YSO.
\end{enumerate}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords }\OperatorTok{=}\NormalTok{ Keywords.extract\_keywords(}\StringTok{"/sample/sample\_master.json"}\NormalTok{)}
\NormalTok{Utility.save\_json(keywords, }\StringTok{"/keywords/keywords\_extracted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_extracted }\OperatorTok{=}\NormalTok{ Utility.load\_json(}\StringTok{"/keywords/keywords\_extracted.json"}\NormalTok{)}
\NormalTok{keywords\_clean }\OperatorTok{=}\NormalTok{ Keywords.clean\_keywords(keywords\_extracted)}
\NormalTok{Utility.save\_json(keywords\_clean, }\StringTok{"keywords/keywords\_clean.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_clean }\OperatorTok{=}\NormalTok{ Utility.load\_json(}\StringTok{"/keywords/keywords\_clean.json"}\NormalTok{)}
\NormalTok{keywords\_histogram }\OperatorTok{=}\NormalTok{ Keywords.make\_histogram(keywords\_clean)}
\NormalTok{Utility.save\_json(keywords\_histogram, }\StringTok{"keywords/keywords\_clean\_histogram.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sample\_data\_set }\OperatorTok{=}\NormalTok{ Utility.load\_json(}\StringTok{"/indexed/indexed\_master.json"}\NormalTok{)}
\NormalTok{keywords\_counted }\OperatorTok{=}\NormalTok{ Keywords.make\_count(sample\_data\_set)}
\NormalTok{Utility.save\_json(keywords\_counted, }\StringTok{"/analysis/keywords\_counted.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The median number of keywords for an item in the sample data set is 6
with an IQR of 4 (min = 0, Q1 = 4, M = 6, Q3 = 8, max = 87). Of course,
items in the sample data set with a number of keywords below the first
and above the third quartile are highly suspect from a qualitative point
of view: when it comes to subject indexing, some terms are required, but
more is usually worse. Items with too few or too many keywords will be
discussed in more detail in section
\protect\hyperlink{assessment}{section ``Assessment''}

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
https://docs.openrefine.org/manual/reconciling) and the tool of choice
for this task is OpenRefine (see
\protect\hyperlink{introduction}{section ``Introduction''}).

In what follows, I will describe how the cleaned keywords were
reconciled with Wikidata and YSO, and which additional steps for
refinement were undertaken. In total, 2'104 data transformation
operations were performed; the complete operation history is available
as \texttt{/keywords/operation\_history.json}.

@@

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API
for details). To generalize, suggestions with types whose concepts are
proper names are usually incorrect. Based on this observation,
suggestions with types such as ``scholarly article,'' ``clinical
trial,'' and ``scientific journal'' were rejected.\footnote{The complete
  list includes ``academic journal,'' ``open access journal,''
  ``thesis,'' ``doctoral thesis,'' and ``natural number.''} Suggestions
types such as ``human,'' ``album,'' and ``commune of France'' were
manually verified (i.e., checked for correctness).\footnote{The complete
  list includes ``film,'' ``musical group,'' ``business,'' ``literary
  work,'' ``television series,'' ``organization,'' ``family name,''
  ``written work,'' ``video game,'' ``single, television series
  episode,'' ``painting,'' , ``city of the United States,''
  ``magazine,'' ``studio album,'' ``year,'' ``nonprofit organization,''
  ``border town,'' ``international organization,'' ``political party,''
  ``software,'' ``song,'' ``website,'' ``article comic strip,''
  ``collection,'' ``commune of Italy,'' ``fictional human,'' ``film,''
  ``government agency,'' ``village,'' ``academic journal article,''
  ``female given name,'' and ``poem.''}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keywords\_reference }\OperatorTok{=}\NormalTok{ Utility.load\_json(}\StringTok{"/keywords/keywords\_reference.json"}\NormalTok{)}
\NormalTok{keywords\_reference\_new }\OperatorTok{=}\NormalTok{ Keywords.enrich\_with\_yso(keywords\_reference)}
\NormalTok{Utility.save\_json(keywords\_histogram, }\StringTok{"keywords/keywords\_reference\_master.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

In this section I will assess the quality of the sample data set's
indexing by Annif based on the gold standard.

@@

Let us briefly look at the definitions and motivations of the chosen
metrics. Remember that a suggestion of a subject term is correct if and
only if the subject term is in the derivative gold standard. The
possible outcomes are summarized in Table 2.

@@

The above scoring metrics were implemented using the scikit-learn
machine learning library (Pedregosa et al. 2011).

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Analysis.super\_make\_metrics(}\StringTok{"/indexed/indexed\_master.json"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}
\centering
\includegraphics{images/metrics_all_n.pdf}
\caption{Distribution of weighted F1-scores per maximum number of
suggestions 1 \(\leq\) n \(\leq\) 10. See Table 3 for numeric values. n
= 2 is the best perfomring type of configuration. However, it is
outperformed by n = 4 and n = 5 token configurations. The distribution
of the distributions is skewed left. Types of configurations with a
smaller maximum number of suggestions display a more consistent
performance.}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{for}\NormalTok{ dept }\KeywordTok{in}\NormalTok{ Data.get\_departments():}
\NormalTok{    Analysis.super\_make\_metrics(}\StringTok{"/indexed/indexed\_master.json"}\NormalTok{, dept)}
\end{Highlighting}
\end{Shaded}
